---
title: "host-phage modeling"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

# Setup Work Environment

```{r}
# Clear and set working directory
rm(list=ls())
getwd()
setwd("~/GitHub/community-modules/")

# Load packages
require("png")
require("grid")
require("deSolve")
require("rootSolve")
require("ggplot2")
require("dplyr")
require("tidyr")
```

# Host-phage model: ecological
```{r}

# Differential equations
HPmod <- function(t, x, parms){
  with(as.list(c(parms, x)), {
    
    dN <- r*N*(1-(N+I)/K) - phi*N*V - omega*N
    dI <- phi*N*N - eta*I - omega*I
    dV <- beta*eta*I - phi*N*V - omega*V
    res <- c(dN, dI, dV)
    return(list(res))
  })
}

# r = host growth rate
# K = carrying capacity
# phi = adsorption rate
# omega = 
# eta = 
# beta = burst size

# Times, intial conditions, and parameters
times <- seq(0, 1000, by = .1)
x0 <- c(N = 10e4, I = 0, V = 1)
parms <- c(eta = 0.01, r = 0.5, K = 10^9, phi = 10e-9, beta = 50, omega = 0.05)

# Solve differential equations and plot output
out <- lsoda(x0, times, HPmod, parms)
plot(out)
plot(out[,"N"], out[,"V"], type = 'l', ylab = "V", xlab = "N", main = "Phase Plane")

# Solve for equilibrium
stode(y = (tail(out, n = 1)[,-1]), func = HPmod, parms = parms, pos = T)
```


bifurc <- list()
eta.vec <- seq(1e-8, 1e6, length.out = 30)
bifurc <- matrix(nrow = length(eta.vec), ncol = length(x0))
for(eta in 1:length(eta.vec)){
  parms <- c(eta = eta.vec[eta], r = 0.16, K = 2.2e7, phi = 10e-9, beta = 50, omega = 0.05)
  out <- lsoda(x0, times, HPmod, parms)
  #bifurc[[paste("eta_",as.character(eta),"_dyn",sep = "")]] <- out
  ss <- stode(y = (tail(out, n = 1)[,-1]), func = HPmod, parms = parms, pos = T)
  #bifurc[[paste("eta_",as.character(eta),"_ss",sep = "")]] <- ss
  if(attributes(x = ss)$steady){
    bifurc[eta,] <- as.vector(ss$y)
  } else bifurc[eta,] <- rep(NA, length(x0))
}
bifurc <- cbind.data.frame(eta.vec, bifurc)
colnames(bifurc) <- c("eta", "N", "I", "V")
bifurc <- gather(bifurc, -eta, key = "Variable", value = "Density")

ggplot(bifurc, aes(x = eta, y = Density, color = Variable)) + 
  geom_line() + 
  scale_x_log10() + 
  scale_y_log10() +
  theme_light()


HPcrypticmod <- function(t, x, parms){
  with(as.list(c(parms, x)), {
    
    dN <- r*N*(1-(N+I+Nr)/K) - phi*N*V - omega*N
    dNr <- rr*Nr*(1-(N+I+Nr)/K) - omega*Nr
    dI <- phi*N*N - eta*I - omega*I
    dV <- beta*eta*I - phi*N*V - omega*V
    res <- c(dN, dNr, dI, dV)
    return(list(res))
  })
}

times <- seq(0, 1000, by = .1)
x0 <- c(N = 10e4, Nr = 1, I = 0, V = 1)
parms <- c(eta = .01, r = 0.5, rr = 0.1, K = 10^7, phi = 10e-9, beta = 50, omega = 0.05)
out <- lsoda(x0, times, HPcrypticmod, parms)
plot(out)
plot(times, out[,"N"] + out[,"Nr"], type = 'l')
plot(out[,"N"], out[,"Nr"], type = 'l', ylab = "Nr", xlab = "N", main = "Phase Plane")

stode(y = (tail(out, n = 1)[,-1]), func = HPcrypticmod, parms = parms, pos = T)

rr.vec <- seq(0.001, 1, length.out = 30)
bifurc <- matrix(nrow = length(rr.vec), ncol = length(x0))
for(rr in 1:length(rr.vec)){
  parms <- c(eta = .01, r = 0.5, rr = rr.vec[rr], K = 10^7, phi = 10e-9, beta = 50, omega = 0.05)
  out <- lsoda(x0, times, HPcrypticmod, parms)
  #bifurc[[paste("eta_",as.character(eta),"_dyn",sep = "")]] <- out
  ss <- stode(y = (tail(out, n = 1)[,-1]), func = HPcrypticmod, parms = parms, pos = T)
  #bifurc[[paste("eta_",as.character(eta),"_ss",sep = "")]] <- ss
  if(attributes(x = ss)$steady){
    bifurc[rr,] <- as.vector(ss$y)
  } else bifurc[rr,] <- rep(NA, length(x0))
}
bifurc <- cbind.data.frame(rr.vec, bifurc)
colnames(bifurc) <- c("rr", "N", "Nr", "I", "V")
bifurc <- gather(bifurc, -rr, key = "Variable", value = "Density")

ggplot(bifurc, aes(x = rr, y = Density, color = Variable)) + 
  geom_line() + 
  theme_light() +
  xlab("Resistant Mutant (Nr) Growth Rate")
